\documentclass[11pt,letterpaper]{article}
\usepackage[letterpaper]{geometry}
\usepackage{acl2012}
\usepackage{times}
\usepackage{latexsym}
\usepackage[fleqn]{amsmath}
\usepackage{multirow}
\usepackage{url}
\makeatletter
\newcommand{\@BIBLABEL}{\@emptybiblabel}
\newcommand{\@emptybiblabel}[1]{}
\makeatother
\usepackage[
            pdfauthor={Edwin Simpson and Iryna Gurevych},
            pdftitle={Finding Convincing Arguments Using Scalable Bayesian Preference Learning},
            pdfsubject={Computer Science},
            pdfkeywords={argument mining, machine learning, NLP, Bayesian methods, variational inference, SVI, convincingness},
            hidelinks,
            bookmarks=false]{hyperref}
\DeclareMathOperator*{\argmax}{arg\,max}
\setlength\titlebox{6.5cm}    % Expanding the titlebox

\usepackage[draft]{todonotes}

\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{algorithm2e}
\usepackage{array}
\usepackage[caption=false,font=footnotesize]{subfig}
\usepackage{tabularx}
\usepackage{numprint}

\newcommand{\bs}{\boldsymbol}  
\newcommand{\wrtd}{\mathrm{d}}

\makeatletter
\makeatother %some sort of hack related to the symbol @

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{ 
Finding Convincing Arguments Using Scalable Bayesian Preference Learning
}

\author{Edwin Simpson and Iryna Gurevych\\
  Ubiquitous Knowledge Processing Lab (UKP)\\
  Department of Computer Science \\
  Technische Universit{\"a}t Darmstadt  \\
  {\tt \{simpson, gurevych\}@ukp.informatik.tu-darmstadt.de} \\}

\begin{document}

\maketitle
\begin{abstract}
We introduce a scalable Bayesian preference learning method for identifying
convincing arguments in the absence of gold-standard ratings or rankings.
In contrast to previous work, we avoid the need for separate methods
to perform quality control on training data, predict rankings and perform pairwise classification.
Bayesian approaches are an effective solution when faced with sparse or noisy training data, 
but have not previously been used to identify convincing arguments.
One issue is scalability, which we address by developing a 
stochastic variational inference method for Gaussian process (GP) preference learning.
We show how our method can be applied to predict argument convincingness from crowdsourced data, 
outperforming the previous state-of-the-art, particularly when trained with small amounts of unreliable data.  
We demonstrate how the Bayesian approach enables more effective active learning,
thereby reducing the amount of data required to identify convincing arguments for new users and domains.
While word embeddings are principally used with neural networks, our results show that word embeddings in combination with linguistic features also benefit GPs when predicting argument convincingness.
\end{abstract}

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
%\IEEEpeerreviewmaketitle    

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{sections/intro}
\input{sections/method}
\input{sections/experiments}
\input{sections/discussion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% use section* for acknowledgement
\section*{Acknowledgements}

This work has been supported by the German Federal Ministry of Education and Research (BMBF) 
under the promotional reference 01UG1416B (CEDIFOR).
It also received funding from the European Union's
Horizon 2020 research and innovation programme (H2020-EINFRA-2014-2) under grant agreement No. 654021 
(Open-MinTeD). It reflects only the authors’ views and the EU is 
not liable for any use that may be made of the information contained therein. 
We would like to thank the TACL editors and reviewers for their effort and the valuable feedback we received from them.

\bibliographystyle{acl2012}
\bibliography{simpson_pref_learning_for_convincingness}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}
